### 2015-06-14 writing/S'15 LSTM/data.tex
Delete SAT instances; logging; paper push.

\section{Recursive structure in artificial data}\label{sec:recursion}
\paragraph{Reasoning about entailment} 
The data that we use define a version of the recognizing textual entailment task, in which the goal is to determine what kind of logical consequence relation holds between two sentences, drawing on a small fixed vocabulary of relations such as entailment, contradiction, or synonymy. This task is well suited to evaluating neural network models for sentence interpretation: models must develop comprehensive represenations of the meanings of each sentence to do well at the task, but the data does not force these representations to take a specific form, allowing the model to learn whatever kind of representations it can use most effectively.

The data we use are labeled with the seven mutually exclusive logical reations of \newcite{maccartney2009extended}, which distinguish entailment in two directions ($\natfor$, $\natrev$), equivalence ($\nateq$), two types of contradiction ($\natneg$, $\natalt$), and two types of semantic independence ($\natind$, $\natcov$).

@@

\paragraph{The artificial language} The language described in Bowman et al.~(\S4) is designed to highlight the use of recursive structure with minimal additional complexity. It's vocabulary consists only of six unanalyzed word types ($p_1, p_2, p_3, p_4, p_5, p_6$), \word{and}, \word{or}, and \word{not}. Sentences of the language can be straightforwardly interpreted as statements of propositional logic (where the six unanalyzed words are variables), and labeled sentence pairs can be interpreted as theorems of that logic. Some example pairs are provided in Table \ref{tab:plexs}.

@@

The data come with parentheses representing complete binary bracketing. Our two models use this information differently. For the tree structured model, we assumble the tree structure for each sentence according to the parentheses, but do not treat them as word tokens. For the sequence model, we treat the parentheses as words, and learn their embeddings. This provides the two models with equivalent data, so that their ability to handle unseen structures can be fairly compared.

\paragraph{The corpus}
The sentence pairs in the corpus are divided into thirteen bins according to the number of logical connectives (\word{and, or, not}) in the longer of the two sentences in the pair. We test the model on each bin separately (58k total examples, using an 80/20\% train/test split) in order to evaluate how its performance depends on the complexity of the sentences. In three experiments, we train our models on the training portions of bins 0--3 (62k examples), 0--4 (90k), and 0--6 (160k). Capping the size of the training sentences below length 12 allows us to evaluate how the models interpret the sentences: if their performance falls off abruptly above the cutoff, it is reasonable to assume that the models are depending heavily on specific sentence structures, and cannot generalize to new structures. If then performance decays gradually\footnote{Since sentences are fixed-dimensional vectors of fixed-precision floating point numbers, the models will inevitably will start to make errors on sentences above some length, and the (necessary) use of L2 regularization exacerbates this for sentences which are longer than those seen in training.} with no such abrupt change, then it must have learned a more generally valid interpretation function for the language which respects its recursive structure.

>>>>

\section{Recursive structure in artificial data}\label{sec:recursion}
\paragraph{Reasoning about entailment} 
The data that we use define a version of the recognizing textual entailment task, in which the goal is to determine what kind of logical consequence relation holds between two sentences, drawing on a small fixed vocabulary of relations such as entailment, contradiction, or synonymy. This task is well suited to evaluating neural network models for sentence interpretation: models must develop comprehensive represenations of the meanings of each sentence to do well at the task, but the data do not force these representations to take a specific form, allowing the model to learn whatever kind of representations it can use most effectively.

The data we use are labeled with the seven mutually exclusive logical reations of \newcite{maccartney2009extended}, which distinguish entailment in two directions ($\natfor$, $\natrev$), equivalence ($\nateq$), exhaustive and non-exhaustive contradiction ($\natneg$, $\natalt$), and two types of semantic independence ($\natind$, $\natcov$).

@@

\paragraph{The artificial language} The language described in Bowman et al.~(\S4) is designed to highlight the use of recursive structure with minimal additional complexity. Its vocabulary consists only of six unanalyzed word types ($p_1, p_2, p_3, p_4, p_5, p_6$), \word{and}, \word{or}, and \word{not}. Sentences of the language can be straightforwardly interpreted as statements of propositional logic (where the six unanalyzed words are variables), and labeled sentence pairs can be interpreted as theorems of that logic. Some example pairs are provided in Table \ref{tab:plexs}.

@@

The data come with parentheses representing complete binary bracketing. Our two models use this information differently. For the tree model, we assumble the tree structure for each sentence according to the parentheses, but do not treat them as word tokens. For the sequence model, we treat the parentheses as words, and learn their embeddings. This provides the two models with equivalent data, so that their ability to handle unseen structures can be fairly compared.

\paragraph{The corpus}
The sentence pairs in the corpus are divided into thirteen bins according to the number of logical connectives (\word{and, or, not}) in the longer of the two sentences in the pair. We test the model on each bin separately (58k total examples, using an 80/20\% train/test split) in order to evaluate how its performance depends on the complexity of the sentences. In three experiments, we train our models on the training portions of bins 0--3 (62k examples), 0--4 (90k), and 0--6 (160k). Capping the size of the training sentences below length 12 allows us to evaluate how the models interpret the sentences: if their performance falls off abruptly above the cutoff, it is reasonable to assume that the models are depending heavily on specific sentence structures, and cannot generalize to new structures. If then performance decays gradually\footnote{Since sentences are fixed-dimensional vectors of fixed-precision floating point numbers, the models will inevitably will start to make errors on sentences above some length, and the use of L2 regularization (which helps overall performance) exacerbates this for sentences which are longer than those seen in training.} with no such abrupt change, then it must have learned a more generally valid interpretation function for the language which respects its recursive structure.

### 2015-06-14 writing/S'15 LSTM/discussion.tex
Delete SAT instances; logging; paper push.

The results are shown in Figure \ref{prop-results}\todo{Next version of figures will have vertical dividing lines at 3,4,6}. All four models were able to largely memorize the training data (TODO, TODO, TODO, and TODO for $\le$6, pooling across bins). In the $\le$6 setting, the three tree models were able to perform at above TODO\% on examples of size $\le$2 with a steady decay in performance continuing through TODO\% at size 12, and similar, if steeper, decays with smaller training sets. All four robustly beat the simple baselines reported in Bowman et al.: the most frequent class occurs just over 50\% of the time, and a summing model that ignores word order does reasonably on the shortest examples but falls below 60\% by bin 4.

@@

\paragraph{Conclusion}

We find that all four models can exploit recursive structure to interpret sentences with complex unseen structures, and that tree structured model's biases allow them to learn to do this more effectively from less data. 
We interpret these results as evidence that the similar performance of tree and sequence models can be attributed to their fundamentally similar representational systems, with both models recognizing compositional structure when it is present, the former doing so more reliably, and the latter benefiting elsewhere---either because of architectural choices that make backpropagation more effective, or because interpreting sentences from left to right allows the model to better simulate other aspects of human language comprehension. We finally suggest that, because of the well supported linguistic claim that the kind of recursive structure that we study here is key to language understanding, there is likely to be value in developing sequence models that can more efficiently exploit this structure without fully sacrificing the flexibility that makes these models succeed.

>>>>

The results are shown in Figure~\ref{prop-results}\todo{Next version of figures will have vertical dividing lines at 3,4,6}. All four models were able to largely memorize the training data (TODO, TODO, TODO, and TODO for $\le$6, pooling across bins). In the $\le$6 setting, the three tree models were able to perform at above TODO\% on examples of size $\le$2 with a steady decay in performance continuing through TODO\% at size 12, and similar, if steeper, decays with smaller training sets. All four robustly beat the simple baselines reported in Bowman et al.: the most frequent class occurs just over 50\% of the time, and a summing model that ignores word order does reasonably on the shortest examples but falls below 60\% by bin 4.

@@

\section{Conclusion}

We find that all four models can exploit recursive structure to interpret sentences with complex unseen structures, and that tree models' biases allow them to learn to do this more effectively from less data. 
We interpret these results as evidence that the similar performance of tree and sequence models can be attributed to their fundamentally similar representational systems, with both models recognizing compositional structure when it is present, the former doing so more reliably, and the latter benefiting elsewhere---either because of architectural choices that make backpropagation more effective, or because interpreting sentences from left to right allows the model to better simulate other aspects of human language comprehension. We finally suggest that, because of the well supported linguistic claim that the kind of recursive structure that we study here is key to language understanding, there is likely to be value in developing sequence models that can more efficiently exploit this structure without fully sacrificing the flexibility that makes these models succeed.
